EinsteinPy: Python for General Relativity\footnote{Released on July, 15th, 2019}

### 2020-01-22 paper.tex
complete symbolic, kind of!

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.3
\documentclass{aastex63}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{listings}  % for code in latex
\usepackage{color}
% \usepackage{caption}

@@

\subsection{COORDINATES}
\subsection{SYMBOLIC}
General Relativity required heavy mathematical formulation to derive and describe various tensors attributing to various features essential for cogent description of a arbitrary space-time. We would describe the various quantities currently supported by this module while maintaining brevity. Each quantity has its own class, even scalars, which are treated as 0\textsuperscript{th} order tensors.

All the mathematical quantities build upon \textit{BaseRelativityTensor} as derived classes, which is discussed in detail in \ref{subsec:BaseRelativityTensor}.

\subsubsection{MetricTensor}\label{subsubsec:metric}
Metric Tensor describes the differential length elements required to measure distance in a curved(also applies to flat) space-time. It is a second order tensor and is fundamental to describe any space-time geometry. The \textit{MetricTensor} class, being inherited from \textit{BaseRelativityTensor} (\ref{subsec:BaseRelativityTensor}) , inherits all its functions and also have some functions and constraints of its own. For example, metric tensor is bound to be a second order tensor.

@@

\subsubsection{ChristoffelSymbols}\label{subsubsec:chl}
\subsubsection{RiemannCurvatureTensor}\label{subsubsec:riemann}
\subsubsection{RicciTensor}\label{subsubsec:riccitensor}
\subsubsection{RicciScalar}\label{subsubsec:ricciscalar}
\subsubsection{Other Defined Quantities}\label{subsubsec:other}

>>>>

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.3
\documentclass{aastex63}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage{listings}  % for code in latex
\usepackage{color}
\usepackage{hyperref}
% \usepackage{caption}

@@

\subsection{COORDINATES}
\subsection{SYMBOLIC}
General Relativity required heavy mathematical formulation to derive and describe various tensors attributing to various features essential for cogent description of an arbitrary space-time. We would describe the various quantities currently supported by this module while maintaining brevity. Each quantity has its own class, even scalars, which are treated as 0\textsuperscript{th} order tensors.

All the mathematical quantities build upon \hyperref[subsec:BaseRelativityTensor]{\textit{BaseRelativityTensor}} as derived classes, which is discussed in detail in \ref{subsec:BaseRelativityTensor}.

\subsubsection{MetricTensor}\label{subsubsec:metric}
Metric Tensor describes the differential length elements required to measure distance in a curved(also applies to flat) space-time. It is a second order tensor and is fundamental to describe any space-time geometry. The \textit{MetricTensor} class, being inherited from \hyperref[subsec:BaseRelativityTensor]{\textit{BaseRelativityTensor}} , inherits all its functions and also have some functions and constraints of its own. For example, metric tensor is bound to be a second order tensor.

@@

\subsubsection{ChristoffelSymbols}\label{subsubsec:chl}

Although Christoffel Symbols are belong to class of pseudo-tensors, it is inherited from \hyperref[subsec:BaseRelativityTensor]{\textit{BaseRelativityTensor}} to maintain homogenity within the module. Being inherited from  \hyperref[subsec:BaseRelativityTensor]{\textit{BaseRelativityTensor}}, it already supports the functions of its parent class and imposes some restrictions of its own, such as order of the tensor should be 3.

Instance of the class can be made by directly passing a 3\textsuperscript{rd} order tensor or can be obtained using the following classmethods,
\begin{itemize}
\item \texttt{from\_metric} : Calculates christoffel symbols from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
\end{itemize}

The defined class function(s) are : 
\begin{itemize}
\item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
\end{itemize}

\subsubsection{RiemannCurvatureTensor}\label{subsubsec:riemann}

This quantity is the most extensive measure of curvature of a sapce-time. It is inherited from \hyperref[subsec:BaseRelativityTensor]{\textit{BaseRelativityTensor}} and supports all the functions of its parent class.

Instance of the class can be made by directly passing a 4\textsuperscript{th} order tensor or can be obtained using the following classmethods
\begin{itemize}
\item \texttt{from\_metric} : Calculates Riemann Tensor from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
\item \texttt{from\_christoffels} : Calculates Riemann Tensor from a given instance of \hyperref[subsubsec:chl]{\textit{ChristoffelSymbols}} class.
\end{itemize}

The defined class function(s) are : 
\begin{itemize}
\item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
\end{itemize}

\subsubsection{RicciTensor}\label{subsubsec:riccitensor}

This tensor is used in solving Eintein's  equation and basically contraction of Riemann Curvature Tensor.  It is inherited from \hyperref[subsec:BaseRelativityTensor]{\textit{BaseRelativityTensor}} and supports all the functions of its parent class.

Instance of the class can be made by directly passing a 2\textsuperscript{nd} order tensor or can be obtained using the following classmethods
\begin{itemize}
\item \texttt{from\_metric} : Calculates Ricci Tensor from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
\item \texttt{from\_christoffels} : Calculates Ricci Tensor from a given instance of \hyperref[subsubsec:chl]{\textit{ChristoffelSymbols}} class.
\item \texttt{from\_riemann} : Calculates Ricci Tensor from a given instance of \hyperref[subsubsec:riemann]{\textit{RiemannCurvatureTensor}} class.
\end{itemize}

The defined class function(s) are : 
\begin{itemize}
\item \texttt{change\_config} : returns a different instance of the same class with different index (contravariant or covariant) configuration. 
\end{itemize}

\subsubsection{RicciScalar}\label{subsubsec:ricciscalar}

It is a scalar quantity obtained by contracting a given Ricci Tensor. It is inherited from \hyperref[subsec:BaseRelativityTensor]{\textit{BaseRelativityTensor}} and supports all the functions of its parent class. Being a scalar quantity, it does not support \texttt{change\_config} unlike other tensors as discussed above.

Instance of the class can be made by directly passing a symbolic expression or can be obtained using the following classmethods
\begin{itemize}
\item \texttt{from\_metric} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:metric]{\textit{MetricTensor}} class.
\item \texttt{from\_christoffels} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:chl]{\textit{ChristoffelSymbols}} class.
\item \texttt{from\_riemann} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:riemann]{\textit{RiemannCurvatureTensor}} class.
\item \texttt{from\_riccitensor} : Calculates Ricci Scalar from a given instance of \hyperref[subsubsec:riccitensor]{\textit{RicciTensor}} class.
\end{itemize}

The defined class properties are : 
\begin{itemize}
\item \texttt{expr} : returns the scalar symbolic expression contained in the class. 
\end{itemize}

Here is a sample code snippet, using the above discussed Classes

\begin{lstlisting}[language=Python, caption=Importing Anti De-Sitter metric]
>>> from einsteinpy.symbolic.predefined import AntiDeSitter
>>> from einsteinpy.symbolic import RicciTensor, RicciScalar
>>> import sympy
\end{lstlisting}

\begin{lstlisting}[language=Python, caption=Obtaining Ricci Tensor from the metric]
>>> RT = RicciTensor.from_metric(AntiDeSitter())
>>> RT.tensor()
\end{lstlisting}

\begin{center}
$\displaystyle \left[\begin{matrix}3 & 0 & 0 & 0\\0 & - 3 \cos^{2}{\left(t \right)} & 0 & 0\\0 & 0 & \left(\sin^{2}{\left(t \right)} - 1\right) \sinh^{2}{\left(\chi \right)} - 2 \cos^{2}{\left(t \right)} \sinh^{2}{\left(\chi \right)} & 0\\0 & 0 & 0 & \left(\sin^{2}{\left(t \right)} - 1\right) \sin^{2}{\left(\theta \right)} \sinh^{2}{\left(\chi \right)} - 2 \sin^{2}{\left(\theta \right)} \cos^{2}{\left(t \right)} \sinh^{2}{\left(\chi \right)}\end{matrix}\right]$
\end{center}

\begin{lstlisting}[language=Python, caption=Deriving Ricci(Curvature) Scalar from Ricci Tensor]
>>> RS = RicciScalar.from_riccitensor(RT)
>>> sympy.simplify(RS.expr)  # simplifying the expression
\end{lstlisting}

\begin{center}
$\displaystyle -12$
\end{center}

In congruence to theoretical results, the code outputs a constant negetive scalar curvature for Anti De-Sitter spacetime. All the above discussed classes are called in this snippet as \hyperref[subsubsec:riccitensor]{\textit{RicciTensor}} internally calls \hyperref[subsubsec:riemann]{\textit{RiemannCurvatureTensor}}, which in turn calls \hyperref[subsubsec:chl]{\textit{ChristoffelSymbols}}.

\subsubsection{Other Defined Quantities}\label{subsubsec:other}

As the library is in heavy development, new quantities are being added as and when required. Some other quantities currently supported are listed below. The quantities are all inherited from \hyperref[subsec:BaseRelativityTensor]{\textit{BaseRelativityTensor}}, and a similar API (same as that of discussed above) has been implemented to maintain compatiblity and homegenity. The class names denote usual quantities relevant to General Relativity.

\begin{itemize}
\item \textit{EinsteinTensor}
\item \textit{StressEnergyMomentumTensor}
\item \textit{WeylTensor}
\item \textit{SchoutenTensor}
\end{itemize}
